---
title: "Data Mung Danielle"
output:
  html_document:
    df_print: paged
---


```{r}
#Install CollegeScorecard Data 
#Load library rscorecard
library(rscorecard)
library(dplyr)
#Obtain API key and load
sc_key('f8fmT8tUvjmgBHGBinSGIS2kQdXIF5Sp4nFp74P0')

```

```{r}
##Extract Institution Name, State Abbr, Average Sat score, InState Tuition, &Pellgrant
# from college in SouthEast region that predominately offer Bachelors

df2014 <- sc_init() %>% 
  sc_filter(region == 5,preddeg == 3) %>% 
  sc_select( instnm, stabbr, sat_avg, tuitionfee_in, pctpell
) %>% 
  sc_year(2014) %>% 
  sc_get(
         debug = FALSE,
         print_key_debug = FALSE,
         return_json = FALSE
  )

df2014
```
```{r}
#2015
df2015 <- sc_init() %>% 
  sc_filter(region == 5,preddeg == 3) %>% 
  sc_select( instnm, stabbr, sat_avg, tuitionfee_in, pctpell
  ) %>% 
  sc_year(2015) %>% 
  sc_get(
    debug = FALSE,
    print_key_debug = FALSE,
    return_json = FALSE
  )
df2015
summary(df2015)
```
```{r}
#2016
df2016 <- sc_init() %>% 
  sc_filter(region == 5,preddeg == 3) %>% 
  sc_select( instnm, stabbr, sat_avg, tuitionfee_in, pctpell
  ) %>% 
  sc_year(2016) %>% 
  sc_get(
    debug = FALSE,
    print_key_debug = FALSE,
    return_json = FALSE
  )
df2016
summary(df2016)
```


```{r}
#2017
df2017 <- sc_init() %>% 
  sc_filter(region == 5,preddeg == 3) %>% 
  sc_select( instnm, stabbr, sat_avg, tuitionfee_in, pctpell
  ) %>% 
  sc_year(2017) %>% 
  sc_get(
    debug = FALSE,
    print_key_debug = FALSE,
    return_json = FALSE
  )
df2017
summary(df2017)
```
```{r}
#2018
df2018 <- sc_init() %>% 
  sc_filter(region == 5,preddeg == 3) %>% 
  sc_select( instnm, stabbr, sat_avg, tuitionfee_in, pctpell
  ) %>% 
  sc_year(2018) %>% 
  sc_get(
    debug = FALSE,
    print_key_debug = FALSE,
    return_json = FALSE
  )

summary(df2018)
```

```{r}
#df18-df16 data frame without Institution/ State Abbr. to plot graphs

df18 <- sc_init() %>% 
  sc_filter(region == 5,preddeg == 3) %>% 
  sc_select(sat_avg, tuitionfee_in, pctpell
  ) %>% 
  sc_year(2018) %>% 
  sc_get(
    debug = FALSE,
    print_key_debug = FALSE,
    return_json = FALSE
  )

df17<- sc_init() %>% 
  sc_filter(region == 5,preddeg == 3 ) %>% 
  sc_select( sat_avg, tuitionfee_in, pctpell
  ) %>% 
  sc_year(2017) %>% 
  sc_get(
    debug = FALSE,
    print_key_debug = FALSE,
    return_json = FALSE
  )
df16<- sc_init() %>% 
  sc_filter(region == 5,preddeg == 3 ) %>% 
  sc_select( sat_avg, tuitionfee_in, pctpell
  ) %>% 
  sc_year(2016) %>% 
  sc_get(
    debug = FALSE,
    print_key_debug = FALSE,
    return_json = FALSE
  )
df15<- sc_init() %>% 
  sc_filter(region == 5,preddeg == 3 ) %>% 
  sc_select( sat_avg, tuitionfee_in, pctpell
  ) %>% 
  sc_year(2015) %>% 
  sc_get(
    debug = FALSE,
    print_key_debug = FALSE,
    return_json = FALSE
  )
df14<- sc_init() %>% 
  sc_filter(region == 5,preddeg == 3 ) %>% 
  sc_select( sat_avg, tuitionfee_in, pctpell
  ) %>% 
  sc_year(2014) %>% 
  sc_get(
    debug = FALSE,
    print_key_debug = FALSE,
    return_json = FALSE
  )




```
```{r}

#install libraries
library(ggplot2)


```

```
```{r}
#Clean data of NAs
attach(df14)
attach(df15)
attach(df16)
attach(rscorecard)
df14n<-df14 %>% dplyr::filter(!is.na(sat_avg), !is.na(pctpell), !is.na(tuitionfee_in))
df14n  
df15n<-df15 %>% dplyr::filter(!is.na(sat_avg), !is.na(pctpell), !is.na(tuitionfee_in))
df15n
df16n<-df16 %>% dplyr::filter(!is.na(sat_avg), !is.na(pctpell), !is.na(tuitionfee_in))
df16n
df17n<-df17 %>% dplyr::filter(!is.na(sat_avg), !is.na(pctpell), !is.na(tuitionfee_in))
df17n 
df18n<-df18 %>% dplyr::filter(!is.na(sat_avg), !is.na(pctpell), !is.na(tuitionfee_in))
summary(df18n)

```



```{r}
#Exploring Sat Score and Percentage of PellGrant Students

df18_pl<-ggplot(df18n , aes(x= sat_avg, y=pctpell))+ geom_jitter(alpha = 0.6)+ stat_smooth(method = "lm", se = F, col="blue")+  ggtitle ("2018 SAT Scores and % of Pell Grant Students")+
  scale_x_continuous("sat_avg", limits = c(700,1550), expand = c(0,0))

df18_pl
```


```{r}
#Condition variable based on Tuition
tuitiongroup <- cut(df14n$tuitionfee_in, c(0,9000,18000,30000,55000))
levels(tuitiongroup) <- c("Cheap" ,"Average","Above Average", "Expensive")
plot(tuitiongroup)

df14_pltuition<-ggplot(df14n , aes(x= sat_avg, y=pctpell, color = tuitiongroup))+ geom_jitter(alpha = 0.6)+ 
    ggtitle ("SAT Scores and % of Pell Grant Students w/ Tuition") + 
    geom_point(data= subset(df14n, tuitiongroup %in% c("Cheap", "Average", "Above Average", "Expensive")))
    
df14_pltuition
```

```{r, include=FALSE}
library(sf)
library(raster)
library(dplyr)
library(spData)
library(tmap) 
library(leaflet)
devtools::install_github("wmurphyrd/fiftystater")
library(fiftystater)

data("fifty_states")
```
``

```{r}
#Pull Median Debt for first Generation Students

Southeastinc <- sc_init() %>% 
  sc_filter(region == 5, preddeg == 3) %>% 
  sc_select( instnm, age_entry, loan_ever  
  ) %>% 
  sc_year(2016) %>% 
  sc_get(
    debug = FALSE,
    print_key_debug = FALSE,
    return_json = FALSE
  )
  
Southeastinc
```



```
```{r}
#Clean NAs
Southeastincn<-Southeastinc %>% dplyr::filter(!is.na(age_entry), !is.na(loan_ever))

Southeastincn

```
```{r}
#subsetdata by age
summary(Southeastincn)
agegroup <- cut(Southeastincn$age_entry, c(18,20,25,28,50))
levels(agegroup) <- c("Teen" ,"Early 20s","Mid20s", "Mature")

ggplot(Southeastincn,
       aes(x = age_entry, y = loan_ever, colour = agegroup)) +
  geom_point(alpha = 0.3,  position = position_jitter()) + stat_smooth(method = "lm") +
  coord_cartesian()

```


```{r}
#showing Overlapping of agegroups
ggplot(Southeastincn, aes(x = age_entry, y = loan_ever, colour = agegroup, shape = as.factor(agegroup))) +
  geom_point(alpha = 0.3,  position = position_jitter())
```

