---
title: "Team Starfish"
output: 
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
    toc_float: yes
---


```{r, eval=TRUE, include=FALSE}

# devtools::install_github("wmurphyrd/fiftystater")
#install.packages("ggrepel")
library(ggrepel)
library(tidyverse)
library(ggplot2)
library(plotly)
library(sf)
library(googlesheets4)
library(fiftystater)

# getwd()
project.dir <- "Homework/Project 1 - College Scoreboard"
dataset.dir <- "Data"
outputs.dir <- "Output"
# setwd(file.path(getwd(), project.dir))
```
   
    
# Import data
```{r}
# Import raw data
raw.df <- read.csv(file = file.path(dataset.dir, "starfishdf.csv"))

# Import field variables
gs4_deauth()
starfish.fields <- read_sheet('https://docs.google.com/spreadsheets/d/1PL5zn6QLU9GSSD8rRreL8r7xaoClrPIrf5qnioZ3eE4/edit?usp=sharing')
```
     
  
# Data Quality   
##### Data Quality Check Function   
This is our function for checking the completeness and free-of-error dimensions of data quality as described by Pipino, Lee, and Wang.   

For completeness, we calculated the ratio of non-na values within each column. So a value of 1 would signify there are no na's in that column.   

For free-of-error, we returned the ratio of outliers within each numeric column. If the column is not numeric, it shows 'not numeric'. Outliers are by default defined as 3IQR less than the 1st quartile, or 3IQR more than the 3rd quartile. The IQR multiplier can be user specified by adding the argument 'criterion= x' where x is the desired multiplier to the function call. So a value of 0 would signify there are no outliers based on the criteria specified.    

```{r}
qualitycheck <- function(df, critereon=3) { 

  # setup
  col_count <- ncol(df)
  
  # return table setup
  returntable <- matrix(nrow=col_count,ncol=2,byrow=TRUE)
  colnames(returntable) <- c("Ratio_Completeness","Ratio_Outliers")
  rownames(returntable) <- colnames(df)
  # returntable <- as.table(returntable)
  
  
  for (i in 1:col_count) {
    
    # Completeness check: percentage of non-na values
    na_count <- sum(length(which(is.na(df[,i]))))
    naratio.count <- (1 - (na_count/nrow(df)) )

    
    # Free-of-error check: count of  outliers
    if ( class(df[1,i])==class(1) || class(df[1,i])==class(1.0) ) {
      IQR <- quantile(df[,i], .75, na.rm = T) - quantile(df[,i], .25, na.rm = T)
      threshold.upper <- quantile(df[,i], .75, na.rm = T) + IQR*critereon
      threshold.lower <- quantile(df[,i], .25, na.rm = T) - IQR*critereon
      outliers.index <- which(df[,i] > threshold.upper | df[,i] < threshold.lower)
      outliers.count <- length(outliers.index)/col_count
    }
    else outliers.count='not numeric'
    
    # Write check values into table:
    returntable[i,] <- matrix(c(naratio.count, outliers.count),ncol=2)
  }
  returndf <- as.data.frame(returntable)
  returndf$Ratio_Completeness <- as.numeric(returndf$Ratio_Completeness)
  
  returndf <- cbind(VAR_NAME = rownames(returndf), returndf)
  rownames(returndf) <- NULL
  
  return(returndf)
  
}
```
      
  
#### Quality check:
```{r}
ordered.df <- raw.df[,order(colnames(raw.df))]

data.quality <- qualitycheck(ordered.df)

hist(data.quality$Ratio_Completeness)
hist(suppressWarnings(as.numeric(data.quality$Ratio_Outliers)))

# data.quality[data.quality$Ratio_Completeness == 0, 1:2]
# data.quality[data.quality$Ratio_Outliers == 'not numeric', ]
```


# Data clean up
Remove all columns that are empty
```{r}
good.fields <- data.quality[data.quality$Ratio_Completeness != 0,1]
clean.df <- ordered.df[, good.fields]
qualitycheck(clean.df)
hist(qualitycheck(clean.df)$Ratio_Completeness)
```



# Total share of enrollment of undergraduate degree-seeking students who are:

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Asian {.tabset}
```{r, include=FALSE}
data("fifty_states")
years = c(2015, 2016, 2017, 2018, "latest")
colors = c("#4CAF50", "#4CAF50", "#4CAF50", "#4CAF50", "#4CAF50")  # in case I want differnt colors later
asians.map <- list()


for (i in 1:5){
  plot.temp <- suppressWarnings(
    ggplot() + geom_polygon(
    data = fifty_states,
    aes(x = long, y = lat, group = group),
    color = "white",
    fill = "grey92") +
    coord_sf(xlim = c(-88, -78),
             ylim = c(24.5, 33),
             expand = FALSE) +
    geom_point(data = clean.df[clean.df$year == years[i],],
               color = colors[i],
               aes(x = longitude,
                   y = latitude,
                   size = ugds_asian)) +
    scale_size(name = "", range = c(.2, 3)) +
    guides(size = guide_legend("asians")) +
    theme_void())
  asians.map[[i]] <- list(plot.temp, years[i])
}
```

### `r asians.map[[1]][[2]]`
```{r, echo=FALSE}
suppressWarnings(ggplotly(asians.map[[1]][[1]]))
```

### `r asians.map[[2]][[2]]`
```{r, echo=FALSE}
suppressWarnings(ggplotly(asians.map[[2]][[1]]))
```     

### `r asians.map[[3]][[2]]`
```{r, echo=FALSE}
suppressWarnings(ggplotly(asians.map[[3]][[1]]))
```    

### `r asians.map[[4]][[2]]`
```{r, echo=FALSE}
suppressWarnings(ggplotly(asians.map[[4]][[1]]))
```    

### `r asians.map[[5]][[2]]`
```{r, echo=FALSE}
suppressWarnings(ggplotly(asians.map[[5]][[1]]))
```    




## Hispanic {.tabset}
```{r, include=FALSE}
data("fifty_states")
years = c(2015, 2016, 2017, 2018, "latest")
dotcolor = "#03A9F4"
peeps.map <- list()


for (i in 1:5){
  plot.temp <- suppressWarnings(
    ggplot() + geom_polygon(
    data = fifty_states,
    aes(x = long, y = lat, group = group),
    color = "white",
    fill = "grey92") +
    coord_sf(xlim = c(-88, -78),
             ylim = c(24.5, 33),
             expand = FALSE) +
    geom_point(data = clean.df[clean.df$year == years[i],],
               color = dotcolor,
               aes(x = longitude,
                   y = latitude,
                   size = ugds_hisp)) +
    scale_size(name = "", range = c(.2, 3)) +
    guides(size = guide_legend("asians")) +
    theme_void())
  peeps.map[[i]] <- list(plot.temp, years[i])
}
```

### `r asians.map[[1]][[2]]`
```{r, echo=FALSE}
suppressWarnings(ggplotly(peeps.map[[1]][[1]]))
```

### `r asians.map[[2]][[2]]`
```{r, echo=FALSE}
suppressWarnings(ggplotly(peeps.map[[2]][[1]]))
```     

### `r asians.map[[3]][[2]]`
```{r, echo=FALSE}
suppressWarnings(ggplotly(peeps.map[[3]][[1]]))
```    

### `r asians.map[[4]][[2]]`
```{r, echo=FALSE}
suppressWarnings(ggplotly(peeps.map[[4]][[1]]))
```    

### `r asians.map[[5]][[2]]`
```{r, echo=FALSE}
suppressWarnings(ggplotly(peeps.map[[5]][[1]]))
```    




## Black {.tabset}
```{r, include=FALSE}
data("fifty_states")
years = c(2015, 2016, 2017, 2018, "latest")
dotcolor = "#673AB7"
peeps.map <- list()


for (i in 1:5){
  plot.temp <- suppressWarnings(
    ggplot() + geom_polygon(
    data = fifty_states,
    aes(x = long, y = lat, group = group),
    color = "white",
    fill = "grey92") +
    coord_sf(xlim = c(-88, -78),
             ylim = c(24.5, 33),
             expand = FALSE) +
    geom_point(data = clean.df[clean.df$year == years[i],],
               color = dotcolor,
               aes(x = longitude,
                   y = latitude,
                   size = ugds_black)) +
    scale_size(name = "", range = c(.2, 3)) +
    guides(size = guide_legend("asians")) +
    theme_void())
  peeps.map[[i]] <- list(plot.temp, years[i])
}
```

### `r asians.map[[1]][[2]]`
```{r, echo=FALSE}
suppressWarnings(ggplotly(peeps.map[[1]][[1]]))
```

### `r asians.map[[2]][[2]]`
```{r, echo=FALSE}
suppressWarnings(ggplotly(peeps.map[[2]][[1]]))
```     

### `r asians.map[[3]][[2]]`
```{r, echo=FALSE}
suppressWarnings(ggplotly(peeps.map[[3]][[1]]))
```    

### `r asians.map[[4]][[2]]`
```{r, echo=FALSE}
suppressWarnings(ggplotly(peeps.map[[4]][[1]]))
```    

### `r asians.map[[5]][[2]]`
```{r, echo=FALSE}
suppressWarnings(ggplotly(peeps.map[[5]][[1]]))
```    




## White {.tabset}
```{r, include=FALSE}
data("fifty_states")
years = c(2015, 2016, 2017, 2018, "latest")
dotcolor = "#FF5722"
peeps.map <- list()


for (i in 1:5){
  plot.temp <- suppressWarnings(
    ggplot() + geom_polygon(
    data = fifty_states,
    aes(x = long, y = lat, group = group),
    color = "white",
    fill = "grey92") +
    coord_sf(xlim = c(-88, -78),
             ylim = c(24.5, 33),
             expand = FALSE) +
    geom_point(data = clean.df[clean.df$year == years[i],],
               color = dotcolor,
               aes(x = longitude,
                   y = latitude,
                   size = ugds_white)) +
    scale_size(name = "", range = c(.2, 3)) +
    guides(size = guide_legend("asians")) +
    theme_void())
  peeps.map[[i]] <- list(plot.temp, years[i])
}
```

### `r asians.map[[1]][[2]]`
```{r, echo=FALSE}
suppressWarnings(ggplotly(peeps.map[[1]][[1]]))
```

### `r asians.map[[2]][[2]]`
```{r, echo=FALSE}
suppressWarnings(ggplotly(peeps.map[[2]][[1]]))
```     

### `r asians.map[[3]][[2]]`
```{r, echo=FALSE}
suppressWarnings(ggplotly(peeps.map[[3]][[1]]))
```    

### `r asians.map[[4]][[2]]`
```{r, echo=FALSE}
suppressWarnings(ggplotly(peeps.map[[4]][[1]]))
```    

### `r asians.map[[5]][[2]]`
```{r, echo=FALSE}
suppressWarnings(ggplotly(peeps.map[[5]][[1]]))
```    




## Women {.tabset}
```{r, include=FALSE}
data("fifty_states")
years = c(2015, 2016, 2017, 2018, "latest")
dotcolor = "#D4E157"
peeps.map <- list()


for (i in 1:5){
  plot.temp <- suppressWarnings(
    ggplot() + geom_polygon(
    data = fifty_states,
    aes(x = long, y = lat, group = group),
    color = "white",
    fill = "grey92") +
    coord_sf(xlim = c(-88, -78),
             ylim = c(24.5, 33),
             expand = FALSE) +
    geom_point(data = clean.df[clean.df$year == years[i],],
               color = dotcolor,
               aes(x = longitude,
                   y = latitude,
                   size = ugds_women)) +
    scale_size(name = "", range = c(.2, 3)) +
    guides(size = guide_legend("asians")) +
    theme_void())
  peeps.map[[i]] <- list(plot.temp, years[i])
}
```

### `r asians.map[[1]][[2]]`
```{r, echo=FALSE}
suppressWarnings(ggplotly(peeps.map[[1]][[1]]))
```

### `r asians.map[[2]][[2]]`
```{r, echo=FALSE}
suppressWarnings(ggplotly(peeps.map[[2]][[1]]))
```     

### `r asians.map[[3]][[2]]`
```{r, echo=FALSE}
suppressWarnings(ggplotly(peeps.map[[3]][[1]]))
```    

### `r asians.map[[4]][[2]]`
```{r, echo=FALSE}
suppressWarnings(ggplotly(peeps.map[[4]][[1]]))
```    

### `r asians.map[[5]][[2]]`
```{r, echo=FALSE}
suppressWarnings(ggplotly(peeps.map[[5]][[1]]))
```    




## Men {.tabset}
```{r, include=FALSE}
data("fifty_states")
years = c(2015, 2016, 2017, 2018, "latest")
dotcolor = "#EC407A"
peeps.map <- list()


for (i in 1:5){
  plot.temp <- suppressWarnings(
    ggplot() + geom_polygon(
    data = fifty_states,
    aes(x = long, y = lat, group = group),
    color = "white",
    fill = "grey92") +
    coord_sf(xlim = c(-88, -78),
             ylim = c(24.5, 33),
             expand = FALSE) +
    geom_point(data = clean.df[clean.df$year == years[i],],
               color = dotcolor,
               aes(x = longitude,
                   y = latitude,
                   size = ugds_men)) +
    geom_point(data = clean.df[clean.df$year == years[i],],
               color = '#D4E157',
               aes(x = longitude,
                   y = latitude,
                   size = ugds_women)) +
    scale_size(name = "", range = c(.2, 3)) +
    guides(size = guide_legend("asians")) +
    theme_void())
  peeps.map[[i]] <- list(plot.temp, years[i])
}
```

### `r asians.map[[1]][[2]]`
```{r, echo=FALSE}
suppressWarnings(ggplotly(peeps.map[[1]][[1]]))
```

### `r asians.map[[2]][[2]]`
```{r, echo=FALSE}
suppressWarnings(ggplotly(peeps.map[[2]][[1]]))
```     

### `r asians.map[[3]][[2]]`
```{r, echo=FALSE}
suppressWarnings(ggplotly(peeps.map[[3]][[1]]))
```    

### `r asians.map[[4]][[2]]`
```{r, echo=FALSE}
suppressWarnings(ggplotly(peeps.map[[4]][[1]]))
```    

### `r asians.map[[5]][[2]]`
```{r, echo=FALSE}
suppressWarnings(ggplotly(peeps.map[[5]][[1]]))
```    


# AMANADS GRAPHS

My graphs look specifically at the state of Florida to compare NCF with the rest of the state: 
```{r, echo=FALSE}
clean.fl <- clean.df %>% filter(stabbr == "FL" & year == "2015")
clean.fl$ncf <- clean.fl$instnm %in% c("New College of Florida")

```
I subsetted the data to reflect this as well as created a dummy variable to denote which college is New College for graph creation.


## First Generation Students vs. Family Income of Florida Universities  in 2015


```{r, echo=FALSE}
plotNCF1 <- ggplot(clean.fl, aes(x = faminc, y = first_gen, color = factor(control))) +
  xlab("Average Family Income") + 
  ylab("Percentage of First Generation Students") + 
  ggtitle("First Generation Students vs. Family Income of Florida Colleges in 2015") +
  geom_point(na.rm = TRUE) +
  geom_text_repel(data = subset(clean.fl, first_gen < 0.15),
                        mapping = aes(label = instnm)) +
  theme(plot.title = element_text(hjust=0.5)) +
  scale_colour_discrete(name = "Ownership", labels = c("Public", "Private Nonprofit", "Private For-Profit"))  
    
plotNCF1

```

In this graph, I wanted to look at the amount of first generation students that a university has and their average family income. There is the obvious negative relationship between the percentage of first gen students and average family income, as first gen students usually come from poorer families. What I am interested is the place where New College sits. Through my experience with the college, everyone came from already highly educated families that were well-off. And it seems that this is the case. New College has the lowest amount of first gen students in the state, and has a rather higher average family inc, the second highest in Public Schools. 

## Percentage of White Undergraduate Students vs. Average Family Income of Florida Universities in 2015  

```{r, echo = FALSE}

plotNCF2 <- ggplot(clean.fl, aes(x = faminc, y = ugds_white, color = factor(control))) +
  xlab("Average Family Income") +
  ylab("Percentage of White Undergraduate Students") + 
  ggtitle("White Undergrads vs. Family Income of Florida Universities in 2015") +
  geom_point(na.rm = TRUE) + 
  scale_colour_discrete(name = "Ownership", labels = c("Public", "Private Nonprofit", "Private For-Profit")) +
  geom_text_repel(data = subset(clean.fl, ncf == "TRUE"),
                        mapping = aes(label = instnm))

plotNCF2

```

In this graph, instead of first gen students, I looked at the percentage of white students. Once again, I wanted to see the relationship between race and income, but how New College fits into the picture. As we can see, New College is fairly white, and once again pretty well-off.

## Average Cost of Attendence / Tuition for All Public Florida Colleges in 2015

```{r, echo = FALSE}

plotNCF3 <- clean.fl %>% filter(preddeg == "3" & distanceonly == "0" & control == "1" & !is.na(costt4_a)) %>% 
  ggplot(aes(x = reorder(instnm, costt4_a), y = costt4_a)) +
  geom_bar(stat = "identity", aes(fill=ncf)) +
  xlab("Average Cost of Attendence") +
  ylab("College") + 
  ggtitle("Average Cost for Public Florida Colleges in 2015") +
  coord_flip() +
  guides(fill=FALSE)

plotNCF3
```
```{r, echo=FALSE}

plt4 <- clean.fl %>% filter(preddeg == "3" & distanceonly == "0" & control == "1" & !is.na(tuitionfee_in)) %>% 
  ggplot(aes(x = reorder(instnm, tuitionfee_in), y = tuitionfee_in)) +
  geom_bar(stat = "identity", aes(fill=ncf)) +
  xlab("In-State Tuition") +
  ylab("College") + 
  ggtitle("In-State Tuition for Public Florida Colleges in 2015") +  
  coord_flip() + 
  guides(fill = FALSE)

plt4


In these graphs I was just curious about the costs of attendence comparing New College with the rest of the Public Schools. I choose just to focus on Public schools as Private schools are going to be more expensive as a given. From these comparisons, we can see that New College is pretty expensive as a public school in Florida. However, the reason why New College has the lowest first gen students may be cost, but could be other things - its worth investigating this lead further. 
